Check model

### 2024-03-11 Code/11_TestModel.Rmd
Final commit
Final edits

---
title: "Create histogram with risk scores"
author: "Carolien C.H.M. Maas"
date: "December 2022"
output: html_document
---

@@

# Test if model is correct
### Check calibration
Intercept of recurrence should be the same as when setting covariates to "zero" values minus the center
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
options(datadist="dd.training")

@@

### Check mean event rates of risk distribution
#### Recurrence
```{r, eval=TRUE}
# h0 for only last imputed data set
f.Rec.basehaz.5 <- survival::basehaz(f.mi.BS.Rec.5,TRUE)
h0.Rec.5 <- f.Rec.basehaz.5$hazard[f.Rec.basehaz.5$time==max(f.Rec.basehaz.5$time[f.Rec.basehaz.5$time<=horizon])]

# MSM.cal.fact for only last imputed data set
lp.5<-predict(f.mi.BS.Rec.5,
              newdata=mice::complete(imputed.training.data,5), 
              type="lp")
f.mi.5 <- rms::cph(S.MSM.5.training~lp.5,y=TRUE,x=TRUE) 
MSM.cal.fact.5 <- stats::coef(f.mi.5)

sum(h.Rec$density*fun.event(lp=lp.points.Rec.5, h0=h0.Rec.5))
# similar to the 5-year survival probability, e.g. kaplan-meier at 5 years, i.e. survfit at 5 years:
SF<-survival::survfit(S.Rec.5.training~1,data=training.data.to.be.imputed)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
# average probability of recurrence
mean(fun.event(lp=lp.Rec, h0=h0.Rec.5))
```

@@

sum(h.Rec$density*fun.event(lp=lp.points.MSM.5, h0=h0.MSM.5))
# equal to the 5-year survival probability, e.g. kaplan-meier at 5 years, i.e. survfit at 5 years:
SF<-survfit(S.MSM.5.training~1,data=training.data.to.be.imputed)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
# average probability of MSM
mean(fun.event(lp=lp.MSM, h0=h0.MSM.5))
```

@@

### Test proportional hazards assumption
```{r, eval=TRUE}
test.prop<-survival::cox.zph(f.mi.full.Rec.5$fits[[1]])
test.prop
plot(test.prop)
```

>>>>

---
title: "Check model"
author: "Carolien C.H.M. Maas"
date: "November 2023"
output: html_document
---

@@

### Check mean event rates of risk distribution
#### Recurrence
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")

# h0 for only last imputed data set
f.Rec.basehaz.5 <- survival::basehaz(f.mi.BS.Rec.5,TRUE)
h0.Rec.5 <- f.Rec.basehaz.5$hazard[f.Rec.basehaz.5$time==max(f.Rec.basehaz.5$time[f.Rec.basehaz.5$time<=horizon])]

# lp for last imputed data set
lp.Rec<-predict(f.mi.BS.Rec.5,
              newdata=mice::complete(imputed.training.data,5), 
              type="lp")

# calibration factor
f.mi.5 <- rms::cph(S.MSM.5.training~lp.Rec,y=TRUE,x=TRUE) 
MSM.cal.fact.5 <- stats::coef(f.mi.5)

# sum(h.Rec$density*fun.event(lp=lp.points.Rec.5, h0=h0.Rec))
# similar to the 5-year survival probability, e.g. kaplan-meier at 5 years, i.e. survfit at 5 years:
SF<-survival::survfit(S.Rec.5.training~1,data=training.data.to.be.imputed)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
# average probability of recurrence
mean(fun.event(lp=lp.Rec, h0=h0.Rec))
```
# Test if model is correct
### Check calibration
Intercept of recurrence should be the same as when setting covariates to "zero" values minus the center
```{r, eval=TRUE}
load("Z:/Project Melanoom/PaperMelanoma/Data/training/training.data.RData")
options(datadist="dd.training")

@@

@@

# sum(h.Rec$density*fun.event(lp=lp.points.MSM.5, h0=h0.MSM.5))
# equal to the 5-year survival probability, e.g. kaplan-meier at 5 years, i.e. survfit at 5 years:
SF<-survfit(S.MSM.5.training~1,data=training.data.to.be.imputed)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
# average probability of MSM
mean(fun.event(lp=MSM.cal.fact*lp.Rec, h0=h0.MSM))
```

@@

### Test proportional hazards assumption
```{r, eval=TRUE}
test.prop<-survival::cox.zph(f.mi.full.Rec.5$fits[[1]])
test.prop
plot(test.prop)
```
### See if RFS always lower than MSS
```{r, eval=TRUE}
p.RFS <- fun.event(lp=predict(f.mi.BS.Rec.5), h0=h0.Rec)
p.MSS <- fun.event(lp=MSM.cal.fact*predict(f.mi.BS.Rec.5), h0=h0.MSM)
plot(p.RFS, p.MSS)
abline(a=0, b=1, col="red")
```
